### 2011-04-27 standing-patches-paper.tex
added examples section

Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
Therefore, if we define $\chi$ to be the unique positive solution to
\[
    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
\]
then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
before it is expected that one other successful mutation would have arisen in the area covered so far.
The explicit formula for $\chi$ is cumbersome; here we omit it.
(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
This can be rewritten 
\begin{equation} \label{eqn:defines_chi}
%    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^3 = (\pi \rho \mu)^{-1} .
\left( \chi s_d \sigma \sqrt{2s_b} \right)^2 + \left( \chi s_d \sigma \sqrt{2s_b} \right)^3 
        = (s_d^3 \sigma^2)/(\pi \rho \mu).
\end{equation}
From this we see that $\chi$ increases {\em linearly} with $1/\sqrt{s_b}$,
and decreases with $\rho$ and $\mu$.
Furthermore, the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
If $\sigma^3 \s_d^3/(\pi \rho \mu)$ is small, then $\chi \approx \sqrt{ s_d / (2 s_b \pi \rho \mu)} = 1/\sqrt{\pi \lambda_0}$,
% + O( \gamma ) with $\gamma = \sigma^3 \s_d^3/(\pi \rho \mu)$
and nearly all mutants come from standing variation (note that $\sigma$ does not enter into it).
On the other hand, if $\sigma^3 \s_d^3/(\pi \rho \mu)$ is large, then $\chi \approx (2s_b)^{-1/2} \left( \sigma \pi \rho \mu)\right)^{-1/3}$,
% + O( 1 ) .
the characteristic length from \citet{ralphcoop2010},
in which case standing variation is relatively unimportant.
% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.

>>>>

Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
Therefore, if we define $\chi$ to be the unique positive solution to
\[
    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
\]
then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
before it is expected that one other successful mutation would have arisen in the area covered so far.
The explicit formula for $\chi$ is cumbersome; here we omit it.
(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
This can be rewritten 
\begin{equation} \label{eqn:defines_chi}
%    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^3 = (\pi \rho \mu)^{-1} .
\left( \chi s_d \sigma \sqrt{2s_b} \right)^2 + \left( \chi s_d \sigma \sqrt{2s_b} \right)^3 
        = \frac{ s_d^3 \sigma^2 }{ \pi \rho \mu } .
\end{equation}
From this we see that $\chi$ increases linearly with $1/\sqrt{s_b}$,
and decreases with $\rho$ and $\mu$.
Furthermore, the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
For instance, if $\sigma^3 s_d^3/(\pi \rho \mu)$ is small, then $\chi \approx \sqrt{ s_d / (2 s_b \pi \rho \mu)} = 1/\sqrt{\pi \lambda_0}$,
% + O( \gamma ) with $\gamma = \sigma^3 s_d^3/(\pi \rho \mu)$
and nearly all mutants come from standing variation.
On the other hand, if $\sigma^3 s_d^3/(\pi \rho \mu)$ is large, then $\chi \approx (2s_b)^{-1/2} \left( \sigma \pi \rho \mu)\right)^{-1/3}$,
% + O( 1 ) .
the characteristic length from \citet{ralphcoop2010},
in which case standing variation is relatively unimportant.
% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.
